% ============================================================
%  Business Incubator Management Platform -- Final Report
%  Software Development Laboratory | IPB-ESTIG | 2025/2026
% ============================================================
\documentclass[12pt,a4paper]{report}

% --- Encoding & Language -------------------------------------
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}

% --- Typography ----------------------------------------------
\usepackage{setspace}
\onehalfspacing

% --- Page Layout ---------------------------------------------
\usepackage[
  top=2.5cm, bottom=2.5cm,
  left=3cm, right=2.5cm,
  headheight=15pt
]{geometry}

% --- Headers & Footers ---------------------------------------
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
\fancyhead[L]{\small\textit{Business Incubator Management Platform}}
\fancyhead[R]{\small\textit{Final Implementation Report}}
\fancyfoot[C]{\thepage}
\renewcommand{\headrulewidth}{0.4pt}

% --- Colours -------------------------------------------------
\usepackage[dvipsnames,svgnames,table]{xcolor}
\definecolor{IPBblue}{RGB}{0,84,166}
\definecolor{IPBgray}{RGB}{88,88,90}
\definecolor{lightgray}{RGB}{245,245,245}
\definecolor{codegray}{RGB}{240,240,240}
\definecolor{successgreen}{RGB}{39,130,67}
\definecolor{warningorange}{RGB}{179,104,0}

% --- Tables --------------------------------------------------
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{longtable}
\usepackage{array}

% --- Lists ---------------------------------------------------

% --- Graphics and Captions -----------------------------------
\usepackage{graphicx}
\usepackage{float}
\usepackage{caption}
\captionsetup{font=small, labelfont=bf, margin=10pt}
\captionsetup[figure]{name=Figure}
\captionsetup[table]{name=Table}

% --- Code blocks ---------------------------------------------
\usepackage{listings}
\lstset{
  backgroundcolor=\color{codegray},
  basicstyle=\ttfamily\small,
  breaklines=true,
  frame=single,
  rulecolor=\color{gray!40},
  numbers=left,
  numberstyle=\tiny\color{gray},
  numbersep=8pt,
  tabsize=2,
  showstringspaces=false,
  captionpos=b
}

% --- Boxes ---------------------------------------------------
\newenvironment{infobox}[1][]{\begin{quote}\noindent\textbf{#1}\par\smallskip}{\end{quote}}
\newenvironment{resultbox}[1][]{\begin{quote}\noindent\textbf{#1}\par\smallskip}{\end{quote}}

% --- Hyperlinks (keep last) ----------------------------------
\usepackage[
  colorlinks=true,
  linkcolor=IPBblue,
  urlcolor=IPBblue,
  citecolor=IPBblue,
  pdftitle={Business Incubator Management Platform - Final Implementation Report},
  pdfauthor={SDL Project Group 20},
  pdfsubject={Software Development Laboratory 2025/2026}
]{hyperref}

% --- Utilities -----------------------------------------------
\usepackage{amsmath}
\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}
\newcolumntype{C}[1]{>{\centering\arraybackslash}p{#1}}
\newcommand{\implemented}{\textcolor{successgreen}{\textbf{Implemented}}}
\newcommand{\partstatus}{\textcolor{warningorange}{\textbf{Partial}}}
\newcommand{\service}[1]{\texttt{#1}}

% ============================================================
\begin{document}

% ============================================================
%  COVER PAGE
% ============================================================
\begin{titlepage}
  \begin{center}
    \vspace*{1.4cm}

    {\Large\bfseries\color{IPBblue}
      Instituto Politécnico de Bragança\\[4pt]
      Escola Superior de Tecnologia e de Gestão}

    \vspace{1.4cm}
    \rule{\linewidth}{1.5pt}
    \vspace{0.5cm}

    {\Huge\bfseries Business Incubator\\[8pt] Management Platform}

    \vspace{0.4cm}
    \rule{\linewidth}{0.5pt}
    \vspace{1.2cm}

    {\Large\bfseries Final Implementation Report}\\[0.5cm]
    {\large Software Development Laboratory | 2025/2026}

    \vfill

    \begin{tabular}{rl}
      \textbf{Course Unit:}   & Software Development Laboratory \\[4pt]
      \textbf{Academic Year:} & 2025/2026 \\[4pt]
      \textbf{Project Group:} & Group 20 \\[4pt]
      \textbf{Submission:}    & Final delivery -- May 2026 \\[4pt]
      \textbf{Defense:}       & Week 15 -- June 2026
    \end{tabular}

    \vspace{2cm}
    {\small\color{IPBgray} Final implementation, deployment, and defense evidence}
  \end{center}
\end{titlepage}

% ============================================================
%  FRONT MATTER
% ============================================================
\pagenumbering{roman}

\chapter*{Abstract}
\addcontentsline{toc}{chapter}{Abstract}

This report presents the final implementation of the \textbf{Business Incubator
Management Platform}, developed for the Software Development Laboratory course
as the continuation of the Phase 1 analysis and design proposal. The original
proposal identified a concrete operational problem: incubators manage resident
companies, occupation contracts, financial obligations, rooms, equipment, and
support requests through fragmented tools. The final system implements that
proposal as a Docker-based microservices platform built primarily with Django
REST Framework, PostgreSQL, RabbitMQ, Nginx, and a Next.js frontend.

The delivered product contains ten bounded backend services, each with its own
runtime responsibility and database boundary, plus a gateway and frontend. It
supports staff operations for company, contract, finance, space, booking,
inventory, ticket, user, dashboard, and document management; client-company
self-service workflows; and public booking requests. The implementation also
includes asynchronous domain events, scheduler and consumer sidecars, seeded
demo data, local verification evidence, a compiled defense report, a presentation
slide deck, and a GitLab CI/CD path for changed-service deployment to the course
production virtual machine.

The report follows the style and structure of the Phase 1 technical document: it
revisits the product context and stakeholder requirements, maps the implemented
system back to the original functional and non-functional requirements, documents
the final architecture and service responsibilities, explains the main
communication flows and deployment workflow, and records the verification and
known operational limitations of the final delivery.

\vspace{1cm}
\noindent\textbf{Keywords:} incubator management, microservices, Django REST
Framework, PostgreSQL, RabbitMQ, Docker Compose, GitLab CI/CD, Next.js, RBAC.

\newpage
\tableofcontents
\newpage
\listoftables
\newpage

% ============================================================
%  MAIN MATTER
% ============================================================
\pagenumbering{arabic}

% ============================================================
\chapter{Introduction and Project Continuity}
\label{ch:introduction}

\section{Purpose of this Report}
\label{sec:purpose}

The first phase of the project produced a formal product and technical proposal
for a Business Incubator Management Platform. That document defined the problem
context, stakeholder goals, functional and non-functional requirements, use
cases, entity model, behavioral sequences, architecture, API catalogue, and user
interface mockups. This final report has a different purpose: it documents the
\textit{implemented result}. It therefore treats the Phase 1 report as a
baseline specification and describes how the delivered software satisfies the
course infrastructure requirements and the original product requirements.

The report is also a defense artifact. It is intended to help evaluators inspect
three dimensions of the delivery:
\begin{enumerate}
  \item \textbf{Product completeness}: whether the implemented platform covers
  the core incubator workflows proposed in Phase 1.
  \item \textbf{Technical compliance}: whether the implementation satisfies the
  Software Development Laboratory Part 2 requirements for microservices,
  databases, RabbitMQ, Docker, reverse proxy, authentication, and deployment.
  \item \textbf{Operational readiness}: whether the repository contains enough
  scripts, documentation, seed data, and verification evidence to run, test, and
  defend the system repeatably.
\end{enumerate}

\section{Original Problem Context}
\label{sec:original-problem}

The Phase 1 report framed the product around a recurring operational problem in
business incubators. These organizations are neither ordinary property managers
nor simple coworking operators. They must monitor resident companies as they
mature, maintain formal occupation contracts, calculate fees, manage physical
spaces and equipment, answer support requests, and provide management reports to
institutional stakeholders. When these activities are handled through
spreadsheets, shared folders, and email, the resulting process becomes fragile:
contract data may diverge from space occupancy data, payment information may not
be visible to resident companies, and management indicators require manual
aggregation.

The implemented platform addresses this problem by consolidating the operational
data into a single system while preserving strict service boundaries. Staff users
receive a management interface for the full incubator portfolio. Client users
receive a portal scoped to their own company. Public users can submit booking
requests without gaining access to private operational data.

\section{Final Delivery Scope}
\label{sec:final-scope}

The final repository delivers a production-shaped application stack with the
following scope:
\begin{itemize}
  \item Ten Django REST Framework backend services: authentication, company,
  contract, finance, space, booking, inventory, ticket, dashboard, and document.
  \item PostgreSQL persistence separated by service databases, matching the
  database-per-service design constraint.
  \item RabbitMQ asynchronous domain events with idempotent consumer handling.
  \item Nginx reverse proxy routing for frontend and API endpoints.
  \item Next.js and Ant Design frontend for staff, client, and public workflows.
  \item MinIO object storage for document payloads, with metadata managed by a
  document service.
  \item Docker Compose runtime with scheduler and consumer sidecars.
  \item GitLab CI/CD build, test, image push, and production deployment workflow.
  \item Defense artifacts: final report, presentation deck, runbooks, local QA
  evidence, and release checklist.
\end{itemize}

\begin{resultbox}[title=Final implementation summary]
The final system is not just a prototype of isolated CRUD screens. It implements
chained business flows across service boundaries: company creation leads to
contracts and documents; contract activation affects spaces and finance;
bookings can reserve spaces and equipment; payments and tickets feed dashboard
indicators; and domain events keep read models and side effects synchronized.
\end{resultbox}

% ============================================================
\chapter{Requirements Baseline}
\label{ch:requirements-baseline}

\section{Course Part 2 Technical Requirements}
\label{sec:course-requirements}

The Software Development Laboratory Part 2 assignment required a microservices
solution implemented with Django REST Framework, PostgreSQL, RabbitMQ, Docker,
Nginx, authentication/authorization, GitLab CI/CD, and documentation. Table
\ref{tab:course-compliance} maps those requirements to the final implementation.

\begin{longtable}{L{0.32\textwidth}L{0.48\textwidth}C{0.13\textwidth}}
\caption{Compliance with Part 2 technical requirements}\label{tab:course-compliance}\\
\toprule
\textbf{Requirement} & \textbf{Implemented evidence} & \textbf{Status} \\
\midrule
\endfirsthead
\toprule
\textbf{Requirement} & \textbf{Implemented evidence} & \textbf{Status} \\
\midrule
\endhead
Minimum four microservices, with at least two core business services & Ten services are implemented. Company, contract, finance, space, booking, inventory, ticket, dashboard, document, and auth are separate Compose services. & \implemented \\
Django REST Framework API backend & Each bounded context is implemented as a Django/DRF service with serializers, viewsets, URLs, migrations, and management commands. & \implemented \\
Minimum eight entities/models & The final domain contains users, companies, employees, contracts, payments, spaces, space types, bookings, booking equipment, equipment types, equipment, tickets, ticket messages, documents, dashboard snapshots, and event-consumption records. & \implemented \\
PostgreSQL persistence & Compose defines service-specific PostgreSQL databases/users, initialized through database scripts and accessed through Django ORM. & \implemented \\
Authentication and authorization & JWT authentication is provided by the auth service; the gateway validates tokens and forwards trusted identity/company headers; services enforce role and company scoping. & \implemented \\
RabbitMQ asynchronous communication & RabbitMQ topic exchange \texttt{incubator.events} carries company, contract, booking, payment, and support events. & \implemented \\
Minimum eight documented events & The repository documents eleven events in \texttt{docs/events.md}. & \implemented \\
Nginx reverse proxy & Gateway routes \texttt{/api/auth/}, \texttt{/api/companies/}, \texttt{/api/contracts/}, etc., and serves the frontend route space. & \implemented \\
Docker Compose orchestration & \texttt{infra/docker-compose.yml} defines infrastructure, backend services, sidecars, frontend, and gateway. & \implemented \\
Separate configuration for environments & Development/runtime configuration is driven by environment files and Compose overlays, including a production image overlay. & \implemented \\
GitLab CI/CD with build, push, deployment stages & The final \texttt{.gitlab-ci.yml} validates, tests, builds changed images, pushes registry tags, and deploys through SSH and Docker Compose. & \implemented \\
Detailed README, report, and presentation material & README, architecture/deploy/event docs, this report, and the PowerPoint deck are included in the repository. & \implemented \\
\bottomrule
\end{longtable}

\section{Phase 1 Product Requirements}
\label{sec:phase1-product-requirements}

The Phase 1 functional catalogue was organized into ten product modules. The
final implementation keeps the same domain decomposition. Table
\ref{tab:functional-traceability} provides a concise traceability view from the
original modules to the delivered functionality.

\begin{longtable}{L{0.22\textwidth}L{0.56\textwidth}C{0.13\textwidth}}
\caption{Traceability from Phase 1 functional modules to implementation}\label{tab:functional-traceability}\\
\toprule
\textbf{Phase 1 module} & \textbf{Final implementation result} & \textbf{Status} \\
\midrule
\endfirsthead
\toprule
\textbf{Phase 1 module} & \textbf{Final implementation result} & \textbf{Status} \\
\midrule
\endhead
Company registration & Company CRUD, employee records, maturity/CAE data, archive state, document links, seeded companies, and staff/client views. & \implemented \\
Company profiles and segmentation & Staff can inspect and edit company data; client portal exposes company-scoped profile data; dashboard includes company/maturity indicators. & \implemented \\
Contract management & Contract CRUD, lifecycle states, company/space association, fee fields, document support, expiration/termination commands, and contract events. & \implemented \\
Financial management & Payments, status tracking, revenue indicators, overdue processing, client payment history, staff finance page, and dashboard graphs. & \implemented \\
Space management & Space types, spaces, bookable flags, price configuration, occupancy/reservation state, contract/booking integration, and public availability. & \implemented \\
Booking system & Staff, client, and public booking flows; staff approval/rejection; reservation availability; optional equipment; price estimation and booking events. & \implemented \\
Inventory and equipment & Equipment types, equipment catalogue, optional space assignment, booking attribution history, rental price metadata, and booking event consumers. & \implemented \\
User management and RBAC & Director/staff/client roles, user CRUD, JWT sessions, company isolation through identity headers, and route guards. & \implemented \\
Support tickets & Client ticket creation, staff queue, messages, status transitions, assignment support, and dashboard ticket indicators. & \implemented \\
Dashboards and reporting & Staff dashboard and finance dashboard include charts, KPIs, revenue/status breakdowns, occupancy summaries, and operational health data. & \implemented \\
\bottomrule
\end{longtable}

\section{Non-Functional Requirement Traceability}
\label{sec:nfr-traceability}

The non-functional requirements from Phase 1 guided implementation decisions.
Security and maintainability received the strongest emphasis because they affect
all services. Table \ref{tab:nfr-traceability} summarizes the main quality
attributes.

\begin{longtable}{L{0.22\textwidth}L{0.58\textwidth}C{0.12\textwidth}}
\caption{Non-functional requirement traceability}\label{tab:nfr-traceability}\\
\toprule
\textbf{Quality attribute} & \textbf{Implementation decision} & \textbf{Status} \\
\midrule
\endfirsthead
\toprule
\textbf{Quality attribute} & \textbf{Implementation decision} & \textbf{Status} \\
\midrule
\endhead
Security & JWT authentication, password hashing through Django, gateway verification, role-aware APIs, company scoping through \texttt{X-Company-Id}, and non-committed deployment secrets. & \implemented \\
Scalability & Service-per-domain Compose containers and database-per-service layout allow independent image builds and service restarts. & \implemented \\
Performance & Dashboard aggregation is delegated to a dashboard service; frontend uses API clients and selected charts; local tests validate expected API helpers. & \implemented \\
Reliability & RabbitMQ consumers store processed \texttt{event\_id} values, scheduler sidecars replace in-process cron, and health checks are declared in Compose. & \implemented \\
Usability & Staff and client shells, dark theme, responsive tables, select-based forms, computed price estimates, and clearer dashboard/finance charts were implemented. & \implemented \\
Maintainability & Monorepo keeps shared infrastructure visible while preserving service-local Django projects, migrations, URLs, and tests. & \implemented \\
Deployment & GitLab CI/CD builds changed images and deploys them through Docker Compose on the assigned VM. & \implemented \\
\bottomrule
\end{longtable}

% ============================================================
\chapter{Final System Architecture}
\label{ch:architecture}

\section{Architectural Pattern}
\label{sec:architectural-pattern}

The final architecture follows the microservice design proposed in Phase 1. Each
bounded context is implemented as a separate Django service, exposed through the
Nginx gateway and deployed as an independent Docker image. The monorepo is used
for course delivery and integration convenience, but the runtime model preserves
service ownership boundaries.

\begin{infobox}[title=Architecture rationale]
The decision to use ten services instead of a smaller four-service design was
intentional. The product domain has naturally separate responsibilities: identity,
company data, contracts, finance, spaces, bookings, inventory, tickets,
dashboard aggregation, and documents. Separating these areas makes the event
flows explicit and supports the course requirement to demonstrate multiple core
business services communicating through APIs and RabbitMQ.
\end{infobox}

\section{Runtime Topology}
\label{sec:runtime-topology}

The runtime topology is composed of four layers:
\begin{enumerate}
  \item \textbf{Edge layer}: Nginx gateway and Next.js frontend.
  \item \textbf{Application layer}: ten Django REST Framework services and
  service-specific scheduler/consumer sidecars.
  \item \textbf{Messaging layer}: RabbitMQ topic exchange for asynchronous
  integration.
  \item \textbf{Persistence layer}: PostgreSQL databases and MinIO object
  storage.
\end{enumerate}

\begin{lstlisting}[caption={Simplified runtime topology}]
Browser
  -> Nginx gateway
     -> Next.js frontend
     -> /api/auth/      -> auth-service      -> auth database
     -> /api/companies/ -> company-service   -> company database
     -> /api/contracts/ -> contract-service  -> contract database
     -> /api/finance/   -> finance-service   -> finance database
     -> /api/spaces/    -> space-service     -> space database
     -> /api/bookings/  -> booking-service   -> booking database
     -> /api/inventory/ -> inventory-service -> inventory database
     -> /api/tickets/   -> ticket-service    -> ticket database
     -> /api/dashboard/ -> dashboard-service -> dashboard database
     -> /api/documents/ -> document-service  -> document database + MinIO

Django services <-> RabbitMQ topic exchange incubator.events
\end{lstlisting}

\section{Service Catalogue}
\label{sec:service-catalogue}

\begin{longtable}{L{0.20\textwidth}L{0.44\textwidth}L{0.20\textwidth}L{0.08\textwidth}}
\caption{Final backend service catalogue}\label{tab:service-catalogue}\\
\toprule
\textbf{Service} & \textbf{Responsibility} & \textbf{Gateway prefix} & \textbf{Port} \\
\midrule
\endfirsthead
\toprule
\textbf{Service} & \textbf{Responsibility} & \textbf{Gateway prefix} & \textbf{Port} \\
\midrule
\endhead
\service{auth-service} & User accounts, JWT login, token verification, role data, and company identity for client users. & \texttt{/api/auth/} & 8001 \\
\service{company-service} & Company profiles, maturity stage, CAE classification, employees, archive status, and company events. & \texttt{/api/companies/} & 8002 \\
\service{contract-service} & Contracts, lifecycle transitions, assigned spaces, rates, monthly fee snapshots, expiration, and contract events. & \texttt{/api/contracts/} & 8003 \\
\service{finance-service} & Payments, billing records, overdue processing, revenue reports, and payment events. & \texttt{/api/finance/} & 8004 \\
\service{space-service} & Space types, spaces, occupancy, reservation blocking, booking availability, and space projections. & \texttt{/api/spaces/} & 8005 \\
\service{booking-service} & Staff/client/public booking requests, approvals, cancellations, quote fields, and booking events. & \texttt{/api/bookings/} & 8006 \\
\service{inventory-service} & Equipment types, equipment items, assignment history, booking-equipment allocation, and inventory consumers. & \texttt{/api/inventory/} & 8007 \\
\service{ticket-service} & Support tickets, messages, assignment/status workflow, and company-scoped ticket access. & \texttt{/api/tickets/} & 8008 \\
\service{dashboard-service} & Aggregated KPIs, service health, ticket metrics, and dashboard read models. & \texttt{/api/dashboard/} & 8009 \\
\service{document-service} & Document metadata, file upload/download/delete, and MinIO object storage integration. & \texttt{/api/documents/} & 8010 \\
\bottomrule
\end{longtable}

\section{Database-per-Service Design}
\label{sec:database-per-service}

The Compose stack creates separate PostgreSQL databases and users for the
services. This directly implements the database-per-service pattern required by
the assignment and proposed in Phase 1. The design avoids shared-table coupling:
services exchange identifiers and selected denormalized fields through APIs or
events rather than joining each other's tables.

The database separation is especially important for:
\begin{itemize}
  \item \textbf{Contract and finance consistency}: contract rates are copied into
  billing/payment contexts rather than recomputed from mutable company data.
  \item \textbf{Space and booking availability}: bookings reserve spaces for time
  windows without making the booking service own the full space catalogue.
  \item \textbf{Client isolation}: downstream services can enforce company scope
  using the trusted identity headers produced by the gateway.
\end{itemize}

\section{Synchronous and Asynchronous Communication}
\label{sec:communication}

The implementation uses synchronous REST calls for user-facing queries and
commands that require immediate responses. It uses RabbitMQ events for
cross-service side effects and eventually consistent read models.

\begin{longtable}{L{0.28\textwidth}L{0.56\textwidth}C{0.10\textwidth}}
\caption{Final domain event catalogue}\label{tab:event-catalogue}\\
\toprule
\textbf{Event} & \textbf{Purpose} & \textbf{Documented} \\
\midrule
\endfirsthead
\toprule
\textbf{Event} & \textbf{Purpose} & \textbf{Documented} \\
\midrule
\endhead
\texttt{company.created} & Notify downstream services that a company profile exists. & Yes \\
\texttt{company.archived} & Preserve historical data while signalling inactive company state. & Yes \\
\texttt{employee.changed} & Update employee-count indicators and company growth data. & Yes \\
\texttt{contract.activated} & Mark contract-dependent projections such as occupied spaces and finance records. & Yes \\
\texttt{contract.expired} & Release dependent resources when contract end date is reached. & Yes \\
\texttt{contract.terminated} & Release dependent resources and preserve termination reason. & Yes \\
\texttt{booking.approved} & Reserve a space/equipment time window and feed dashboards. & Yes \\
\texttt{booking.rejected} & Record staff decision for client/public visibility. & Yes \\
\texttt{booking.cancelled} & Release reservations and equipment allocations. & Yes \\
\texttt{booking.completed} & Mark the end of temporary resource usage. & Yes \\
\texttt{payment.recorded} & Update finance and dashboard indicators after payment. & Yes \\
\bottomrule
\end{longtable}

Every event envelope includes an \texttt{event\_id}. Consumers record processed
IDs so retries do not duplicate side effects. This is an important reliability
property for RabbitMQ because the broker may redeliver messages after network or
consumer failures.

% ============================================================
\chapter{Implemented Product Capabilities}
\label{ch:capabilities}

\section{Staff Back Office}
\label{sec:staff-backoffice}

The staff interface is the operational center of the platform. It exposes the
main administrative domains from the Phase 1 sitemap: dashboard, companies,
contracts, spaces, bookings, inventory, finance, tickets, and users. The UI was
refined beyond placeholder pages into a consistent dark-themed management
experience with navigation, tables, forms, charts, and contextual detail areas.

\subsection{Dashboard and Reporting}
\label{subsec:dashboard-reporting}

The dashboard implements the reporting objectives from FR-1001 and FR-1002. It
shows company totals, maturity/sector breakdowns, occupancy indicators,
financial summaries, ticket metrics, and service health. The finance page adds
charts for revenue and payment state so that staff can quickly understand the
current operational picture without exporting data.

\subsection{Company and Contract Administration}
\label{subsec:company-contract}

Company administration supports legal/profile fields, CAE and maturity data,
employees, and document references. Contract administration supports company and
space association, rate fields, lifecycle state, and document attachment. The
contract lifecycle preserves the Phase 1 rule that active contract economic
values are treated as historical snapshots rather than being silently changed by
future company classification changes.

\subsection{Space and Booking Operations}
\label{subsec:space-booking}

The space module supports space types, space prices, capacity, bookable flags,
occupancy state, and reservation blocking. The booking module supports staff
review of client and public booking requests, approval/rejection workflows, and
booking cost fields. Several UX corrections were made in the final stage:
reservation-related forms use selects and contextual labels instead of raw UUID
text inputs; pricing is computed from space/equipment and duration where
appropriate; and the staff reservation table places detailed information in the
expanded row instead of overloading columns.

\subsection{Inventory, Tickets, and Users}
\label{subsec:inventory-tickets-users}

Inventory management covers equipment types, equipment items, flexible rental
cost metadata, optional assignment to a space, and attribution history. Ticket
management supports client-created support requests, staff status handling, and
message threads. User management supports director/staff/client roles, client
company association, and deactivation/edit flows.

\section{Client Portal}
\label{sec:client-portal}

The client portal implements the Phase 1 requirement that resident companies
should see their own information without accessing other companies' data. Client
routes expose company profile information, contract details, payments, bookings,
and support tickets. The gateway and API clients carry the company identity so
downstream services can enforce \texttt{X-Company-Id} isolation.

The client booking flow was refined to improve usability: available spaces are
loaded into select inputs, equipment options show price context, reserved time
windows are blocked, and estimated price is calculated instead of being entered
as a free-form number. This aligns the implemented product with the original
usability requirements for clear form validation and low-error workflows.

\section{Public Booking Request}
\label{sec:public-booking}

The public booking route allows external users to request bookable spaces by
providing contact and company information. The request enters the booking
workflow as pending and requires staff review, satisfying the Phase 1 public
booking requirement while preserving staff control over actual resource
reservation.

\section{Document Management}
\label{sec:document-management}

The document service manages metadata and MinIO object storage for company and
contract documents. Upload, list, download, and delete flows are available
through service endpoints and integrated with seeded demo data. This supports the
original requirement that signed contracts and supporting documents remain linked
to the appropriate company or contract records.

% ============================================================
\chapter{Implementation Details}
\label{ch:implementation-details}

\section{Repository Structure}
\label{sec:repo-structure}

The repository is organized as a deployable monorepo. The most important paths
are:
\begin{itemize}
  \item \texttt{services/*-service/}: Django REST Framework services.
  \item \texttt{frontend/}: Next.js 14 frontend and API client modules.
  \item \texttt{gateway/}: Nginx routing configuration.
  \item \texttt{infra/docker-compose.yml}: main runtime stack.
  \item \texttt{infra/docker-compose.production.yml}: production image overlay.
  \item \texttt{infra/scripts/ci/}: changed-service detection, image build, and
  remote deploy scripts.
  \item \texttt{libs/py-common/}: shared Python utilities, including event
  envelope support.
  \item \texttt{docs/}: architecture, events, deployment, defense, and report
  artifacts.
\end{itemize}

\section{Authentication and Authorization}
\label{sec:auth-authorization}

Authentication uses email/password login with JWT issuance by the auth service.
The gateway validates incoming tokens and forwards trusted headers to backend
services. This keeps token verification consistent at the edge while allowing
individual services to enforce domain-specific authorization rules. Staff and
director users can access portfolio-wide data; client users are restricted to
their associated company.

\section{Seed Data and Demo Readiness}
\label{sec:seed-data}

The final delivery includes seed data for the primary workflows: users,
companies, contracts, payments, spaces, space types, bookings, inventory items,
tickets, and documents. The seed set is designed for demonstration rather than
for artificial benchmark volume. Its purpose is to ensure that each screen and
business flow has linked data: companies have contracts, contracts have finance
records, spaces and equipment have prices, clients have tickets/bookings, and
documents can be downloaded.

\section{Background Processing}
\label{sec:background-processing}

The implementation intentionally avoids Celery, Kubernetes, and cloud services.
Background work is performed through scheduler and consumer sidecars:
\begin{itemize}
  \item Contract scheduler: detects expirations and publishes lifecycle events.
  \item Finance scheduler: generates/updates billing and overdue payment state.
  \item Booking scheduler: completes or releases time-bound reservations.
  \item Service consumers: process RabbitMQ events and update local projections.
\end{itemize}

This design is simpler than a full distributed job platform but sufficient for
the course requirements and easier to demonstrate through Docker Compose.

\section{Frontend Design and User Experience}
\label{sec:frontend-design}

The Phase 1 report contained mockups for staff and client screens. The final
frontend follows the same navigation concept but adapts the visual design into a
cohesive dark operational theme. Ant Design components provide tables, forms,
selects, date/time inputs, and feedback messages. The final UI work focused on
removing low-quality interactions such as raw UUID inputs, overly wide tables,
misaligned dashboard cards, and missing price context.

The most important UX decisions were:
\begin{itemize}
  \item Use entity labels and selects instead of requiring users to type
  identifiers.
  \item Keep staff tables readable by moving secondary details into expanded
  cards.
  \item Compute estimates from price and duration instead of accepting arbitrary
  values from client users.
  \item Keep the client portal visibly separated from staff administration.
  \item Make dashboard cards and charts visually consistent for defense use.
\end{itemize}

% ============================================================
\chapter{Deployment and CI/CD}
\label{ch:deployment}

\section{Local Runtime}
\label{sec:local-runtime}

The project can be started locally either through the repository Makefile or
through Tilt. Both paths use the same Docker Compose topology: PostgreSQL,
RabbitMQ, Redis, MinIO, backend services, scheduler/consumer sidecars, frontend,
and gateway. Health checks and entrypoint scripts prepare migrations and service
readiness.

The Makefile path is useful for a compact full-stack run and for final demo
validation:

\begin{lstlisting}[language=bash,caption={Make-based local runtime commands}]
cp .env.example .env
make up
make seed
make demo
make logs
make down
\end{lstlisting}

Tilt is useful during development because it watches service-specific folders and
rebuilds or restarts only the affected resources instead of restarting the whole
stack for unrelated file changes. This is especially important in the monorepo,
where documentation, frontend, and backend files are edited frequently.

\begin{lstlisting}[language=bash,caption={Tilt-based local development commands}]
cp .env.example .env
tilt up
# open the Tilt web UI, inspect resource logs, and trigger targeted rebuilds
tilt down
\end{lstlisting}

\section{Production Deployment Workflow}
\label{sec:production-deploy}

The final CI/CD implementation uses GitLab stages for validation, testing,
image build, and production deployment. Changed-service detection reduces the
amount of work performed for each main-branch update. A service-local change
builds and deploys the affected image and any associated sidecar service. Shared
library, Compose, deployment, or environment-template changes trigger a full
rollout.

\begin{lstlisting}[language=bash,caption={Changed-service deployment shape}]
# detect changed deployable services
bash infra/scripts/ci/detect-changed-services.sh

# build and push only affected images
bash infra/scripts/ci/build-and-push.sh

# deploy affected compose services remotely
bash infra/scripts/ci/deploy-remote.sh
\end{lstlisting}

The production deployment target is the assigned course virtual machine. The
user is fixed as \texttt{a69603}; the host and password are stored as protected
GitLab variables. The repository does not contain the VM password, production
environment file, or registry credentials.

\section{Remote Host Preparation}
\label{sec:remote-host}

The course VM was checked over SSH. Docker Engine and Docker Compose v2 were
installed, the service user was added to the \texttt{docker} group, and the
versions were verified. This prepares the VM for the GitLab deployment job once
the pushed pipeline reaches the deploy stage.

\section{Environment Variables and Secrets}
\label{sec:secrets}

The final deployment workflow distinguishes repository configuration from
secret configuration. The repository contains \texttt{.env.example} and CI
scripts, but runtime secrets are provided through GitLab protected variables or
the remote shared \texttt{.env} file. This matches the Part 2 requirement to use
environment variables for credentials and avoids committing secrets to the
repository.

% ============================================================
\chapter{Verification and Evidence}
\label{ch:verification}

\section{Local Verification Strategy}
\label{sec:verification-strategy}

Because the project uses shared GitLab runners, local verification was treated as
the primary merge gate. The goal was to prove correctness before relying on
remote CI availability. Verification was split into host-only checks, frontend
checks, Python checks, Compose checks, report/deck generation, SSH connectivity,
and VM readiness.

\section{Executed Checks}
\label{sec:executed-checks}

\begin{longtable}{L{0.42\textwidth}L{0.42\textwidth}C{0.10\textwidth}}
\caption{Final verification evidence}\label{tab:verification}\\
\toprule
\textbf{Check} & \textbf{Purpose} & \textbf{Result} \\
\midrule
\endfirsthead
\toprule
\textbf{Check} & \textbf{Purpose} & \textbf{Result} \\
\midrule
\endhead
\texttt{bash -n infra/scripts/ci/*.sh} & Validate CI helper shell syntax. & Pass \\
GitLab YAML parse & Validate parent and child pipeline YAML files. & Pass \\
Changed-service detection smoke & Confirm correct deploy variable generation. & Pass \\
Docker Compose production config & Validate base Compose plus production image overlay. & Pass \\
\texttt{npm run lint} & Run root/frontend lint workflow. & Pass \\
\texttt{npm --prefix frontend run typecheck} & TypeScript validation. & Pass \\
\texttt{npm --prefix frontend test} & Frontend unit/API helper tests. & Pass \\
\texttt{npm --prefix frontend run build} & Production frontend build. & Pass \\
\texttt{python3 -m ruff check .} & Python style/static validation. & Pass \\
\texttt{pytest infra/tests} & Infrastructure/runtime wiring tests. & Pass \\
\texttt{latexmk -pdf ... final-report.tex} & Compile this report. & Pass \\
PowerPoint ZIP/XML validation & Confirm deck is a valid PPTX package. & Pass \\
SSH smoke to course VM & Confirm VM login and host identity. & Pass \\
Docker/Compose VM check & Confirm remote Docker Engine and Compose availability. & Pass \\
\bottomrule
\end{longtable}

% ============================================================
\chapter{Limitations and Future Work}
\label{ch:limitations}

\section{Known Limitations}
\label{sec:known-limitations}

The final implementation satisfies the main product and technical requirements,
but several areas remain natural candidates for future improvement:
\begin{itemize}
  \item \textbf{Full localization}: user-facing Portuguese copy is present in the
  UI, but a complete bilingual translation audit would be separate work.
  \item \textbf{Advanced analytics}: dashboard and finance charts cover the main
  defense indicators, while deeper historical reports could be extended with
  exportable filters and longer time-series analysis.
  \item \textbf{Hard production security}: the course deployment uses the assigned
  VM and protected CI variables. A real production installation should add TLS,
  backup automation, external monitoring, and a stricter secrets manager.
  \item \textbf{Performance benchmarking}: functional and build checks were run;
  formal load tests against hundreds of companies remain future work.
\end{itemize}

\section{Future Enhancements}
\label{sec:future-enhancements}

Future work should prioritize the following improvements:
\begin{enumerate}
  \item Add formal OpenAPI publication for every service and link it from the
  README.
  \item Add scheduled backups for PostgreSQL and MinIO on the production VM.
  \item Extend dashboard reporting with date-range filters and downloadable CSV
  summaries.
  \item Add end-to-end Playwright tests for the complete staff/client/public demo
  path once the production deployment is stable.
  \item Introduce monitoring dashboards for RabbitMQ queues, service health, and
  failed background jobs.
\end{enumerate}

% ============================================================
\chapter{Conclusion}
\label{ch:conclusion}

The final implementation converts the Phase 1 proposal into a functioning
microservices platform. It preserves the original product intent---centralized,
role-aware management of incubated companies, contracts, finance, spaces,
bookings, inventory, tickets, documents, and dashboards---while meeting the
technical constraints of Part 2: Django REST Framework services, PostgreSQL,
RabbitMQ, Docker Compose, Nginx reverse proxy, authentication/authorization,
GitLab CI/CD, and deployment automation.

The most important result is the integration between business workflows. The
system does not treat domains as unrelated CRUD tables: contracts influence
spaces and finance; bookings influence spaces and equipment; payments and
tickets feed reporting; documents are attached to company and contract records;
and client-company users are isolated from other companies' data. This is the
core value promised by the original proposal and delivered in the final branch.

Together with the README, architecture documentation, deployment notes, event
catalogue, seed data, presentation deck, and local verification evidence, this
report provides the final technical handoff for evaluation and defense.

\end{document}
